---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  fig.dev = "grDevices::png",
  dpi = 300L,
  dev.args = list(),
  fig.ext = "png",
  fig.retina = 2L,
  fig.align = "center"
)
```

# clusterlayouts

Graph layouts using cluster information. The function calculates a grouped layout based on the node attribute in `igraph` or `tbl_graph` object, while respecting the original edge attributes. If you performed community detection based on the network, these layouts should not be used as this layout just explicitly places nodes belonging to the same category together. The `layout_cluster_wise` uses `igraph::merge_coords` internally to place nodes using the dla algorithm. Reference to `Group Attributes Layout` in Cytoscape and `component_wise` and `layout_components` for the disconnected graphs in `igraph`.

```{r, warning=FALSE, message=FALSE, fig.height=4, fig.width=5}
## Make an example graph and assign a random category
library(clusterlayouts)
library(igraph)
test <- igraph::random.graph.game(n=100, p=0.01)
V(test)$category <- as.factor(sample(1:4, 100, replace=TRUE))
lyt <- clusterlayouts::layout_cluster_panel(test, "category")
plot(test, layout=as.matrix(lyt[,1:2]),
     vertex.color=as.numeric(V(test)$category),
     vertex.size=5, vertex.label=NA)
```

```{r, fig.width=12, fig.height=10, warning=FALSE}
## Can change per cluster layout
test <- igraph::random.graph.game(n=500, p=0.001)
panel <- c(2,3,5,6,6,3,2,2)
V(test)$category <- as.factor(sample(1:sum(panel), 500, replace=TRUE))
lyt <- clusterlayouts::layout_cluster_panel(test, "category", per_row=panel, per_layout="drl")
plot(test, layout=as.matrix(lyt[,1:2]),
     vertex.color=as.numeric(V(test)$category),
     vertex.size=3, vertex.label=NA)
```


```{r, fig.width=12, fig.height=10, warning=FALSE}
## Based on the column
test <- igraph::random.graph.game(n=500, p=0.001)
panel <- c(2,3,5,6,6,3,2,2)
V(test)$category <- as.factor(sample(1:sum(panel), 500, replace=TRUE))
lyt <- clusterlayouts::layout_cluster_panel_col(test, "category", per_col=panel)
plot(test, layout=as.matrix(lyt[,1:2]),
     vertex.color=as.numeric(V(test)$category),
     vertex.size=3, vertex.label=NA)
```

```{r, fig.width=15, fig.height=10, warning=FALSE}
## Can change widths per column
test <- igraph::random.graph.game(n=1000, p=0.001)
panel <- c(2,4,3,1,5,3,6,7,2,4)
widths <- c(1,2,3,4,5,12,3,2,1,2)
V(test)$category <- as.factor(sample(1:sum(panel), 1000, replace=TRUE))
lyt <- layout_cluster_panel_col(test, "category", 
                          per_col=panel, widths=widths)
plot(test, layout=as.matrix(lyt[,1:2]),
     vertex.color=as.numeric(V(test)$category),
     vertex.size=1, vertex.label=NA)
```



